An efficient composition of bidirectional programs by tupling and lazy updates

### 2019-11-29 paper/1.introduction.tex
revising the introduction (on the way)

%In software, there are strong demands for synchronizing data. In database community this is known as ``the view update problem'' and researched for a long time.
% A Survey to View Update Problem
%As a solution for this problem, bidirectional transformation (BX) is introduced.
The synchronization of data is a common problem. In the database community this problem is known as “the view update problem” and has been investigated for a long time~\cite{Bancilhon:1981:USR:319628.319634}. Bidirectional transformation (BX) provides a systematic approach to solving this problem.
As an example, consider a small BX program of $phead$\footnote{The actual program is shown in the next section.}, which consists of two functions: $get$ (for getting the head of an input list) and $put$ (for relfecting the output to the head of the input). Figure \ref{fig:eval-phead} shows the behaviors of these functions when evaluating $phead$.

@@

Here, let $[1,2]$ be the original source \texttt{s}.
$get$ is a projection: $get$ of $phead$ picks the first element of the given original source as a view.
%\texttt{get pHead [1,2] = 1}.
After the view \texttt{v} is obtained, the user can modify the view.
In this case, the user
modified the view
from $1$ to $100$.
From this updated view, how can the source be updated?
This is ``the view update problem.''
To solve this problem, BX provides $put$, an update function on the original source.
In this example, $put$ of $phead$ constructs a new list ${[100,2]}$ from the updated view (\texttt{v'}) and the original source (\texttt{s}).
Note that the update to \texttt{s'} uses \texttt{s} and \texttt{v'} as input.
%\texttt{put pHead [1,2] 8 = [8,2]}

The composition of BX programs is a fundamental construct to build more complex BX programs.
%In BX programs, composition of programs is essential for writing various interesting evaluations.
Because recursions are achieved by compositions, large number of programs include compositions.
In the BX lens papers \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487} compositions are also known as central part of BX programs.
In this section we use a small BX program that contains compositions as a running example, because the programs including recursions can be complicated easily. The running examples are two versions of two composition of $phead$: ($phead \circ phead) \circ phead$ (we call lp3) and $phead \circ (phead \circ phead$) (we call rp3).

@@

Figure \ref{fig:eval-comp-phead} illustrates the evaluation of these programs where the original source \texttt{s} is ${[[[1,2],[3]],[4]]}$ and the updated view is ${100}$.
By repetitious evaluation of $get$s of $phead$, we can obtain the view $v$, ${1}$. To obtain the final updated source \texttt{s'}, we need to evaluate $put$ three times. The first put evaluation is from \texttt{i$_2$} and \texttt{v'} and we can obtain \texttt{i$_2$'}.

First, let us see the standard evaluation method of compositions: ``not keeping any intermediate states and obtaining them by evaluation when they are needed.'' In this example, ``intermediate states'' are \texttt{i$_1$} and \texttt{i$_2$}. This method is used in a BX language, BiGUL \cite{Ko:2016:BFV:2847538.2847544,Ko:2017:ABB:3177123.3158129}. The merit of this method is clear semantics and easy to implement. The disadvantage is the number of $get$s will be quadratic when the BX programs' compositions are left associative. In the evaluation of lp3, two $get$s are required for obtaining \texttt{i$_2$} and one $get$ is required for obtaining \texttt{i$_1$}. In total, three $get$s are evaluated. We explain this method in Section~\ref{sect:minbigul}.

>>>>

%In software, there are strong demands for synchronizing data. In database community this is known as ``the view update problem'' and researched for a long time.
% A Survey to View Update Problem
%As a solution for this problem, bidirectional transformation (BX) is introduced.
The synchronization of data is a common problem. In the database community this problem is known as “the view update problem” and has been investigated for a long time~\cite{Bancilhon:1981:USR:319628.319634}. Bidirectional transformation (BX) provides a systematic approach to solving this problem.
Consider a small BX program of $phead$\footnote{The actual program is shown in the next section.}, which consists of two functions: $get$ (for getting the head of an input list) and $put$ (for reflecting the output to the head of the input). Figure \ref{fig:eval-phead} shows an example of the bidirectional behavior of $phead$.
Let $[1,2]$ be the original source \texttt{s}.
$get$ is a projection: $get$ of $phead$ picks the first element of the given original source $[1,2]$ and returns $1$ as a view $v$.
%\texttt{get pHead [1,2] = 1}.
%After the view \texttt{v} is obtained, the user can modify the view.
%In this case, the user
%modified the view
%from $1$ to $100$.
%From this updated view, how can the source be updated?
%This is ``the view update problem.''
%To solve this problem, BX provides $put$, an update function on the original source.
%In this example,
Supposing that the view is updated to $100$,
$put$ of $phead$ will construct a new source $s'$ of ${[100,2]}$ from the updated view (\texttt{v'}) of $2$ and the original source (\texttt{s}) of $[1,2]$.
%Note that the update to \texttt{s'} uses \texttt{s} and \texttt{v'} as input.
%\texttt{put pHead [1,2] 8 = [8,2]}

@@

The composition of BX programs is a fundamental construct to build more complex BX programs \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487}. Let $bx_1$ (defined by $get_{bx_1}$ and $put_{bx_1}$) and $bx_2$ (defined by $get_{bx_2}$ and $put_{bx_2}$) be two bidirectional programs, then their composition $bx_2 \circ bx_1$ is defined by
\begin{align}
get_{bx_2 \circ bx_1}~s &= get_{bx_2} (get_{bx_1}~s)\\
put_{bx_2 \circ bx_1}~(s,v') &= put_{bx_1} (s,put_{bx_2} (get_{bx_1}~s,v'))
\end{align}

%In BX programs, composition of programs is essential for writing various interesting evaluations.
%Because recursions are achieved by compositions, large number of programs include compositions.
%In the BX lens papers \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487} compositions are also known as central part of BX programs.
%In this section we use a small BX program that contains compositions as a running example, because the programs including recursions can be complicated easily. The running examples are two versions of two composition of $phead$: ($phead \circ phead) \circ phead$ (we call lp3) and $phead \circ (phead \circ phead$) (we call rp3).

@@

One interesting thing is that $put_{bx_2 \circ bx_1}$ needs to call $get_{bx_1}$ to compute the intermediate result, which would introduce an efficient problem when we compose many bidirectional programs. Generally, if we compose $O(n)$ bidirectional programs, we would call $get$ for $O(n^2)$ times. Figure \ref{fig:eval-comp-phead} illustrates the evaluation of our running example:
\[
lp3 = (phead \circ phead) \circ phead
\]
where the original source \texttt{s} is ${[[[1,2],[3]],[4]]}$ and the updated view is ${100}$.
%By repetitious evaluation of $get$s of $phead$, we can obtain the view $v$, ${1}$. To obtain the final updated source \texttt{s'}, we need to evaluate $put$ three times. The first put evaluation is from \texttt{i$_2$} and \texttt{v'} and we can obtain \texttt{i$_2$'}.

%First, let us see the standard evaluation method of compositions: ``not keeping any intermediate states and obtaining them by evaluation when they are needed.'' In this example, ``intermediate states'' are \texttt{i$_1$} and \texttt{i$_2$}. This method is used in a BX language, BiGUL \cite{Ko:2016:BFV:2847538.2847544,Ko:2017:ABB:3177123.3158129}. The merit of this method is clear semantics and easy to implement. The disadvantage is the number of $get$s will be quadratic when the BX programs' compositions are left associative. In the evaluation of lp3, two $get$s are required for obtaining \texttt{i$_2$} and one $get$ is required for obtaining \texttt{i$_1$}. In total, three $get$s are evaluated. We explain this method in Section~\ref{sect:minbigul}.

\textcolor{red}{I am going to revise the rest of the introduction from here.}
